### 2012-11-24 paper.tex
note on validation

We thus inferred the prediction attribute ``alive-at-1'' as explained in the description of the whole dataset: For all patients where the months of survival are greater than or equal to 12 we set the ``alive-at-1''-attribute to 1, otherwise we set it to 0.

The only thing which remained then was to a

\todo{Compare results with/without preprocessing}

@@

\begin{lstlisting}
=== Confusion Matrix ===
  a  b   <-- classified as
  0  4 |  a = 0
  1 90 |  b = 1

@@

pericardial-effusion = 0
|   epss <= 13100: 1 (76.62/1.95)
|   epss > 13100
|   |   epss <= 15600: 0 (2.19/0.16)
|   |   epss > 15600: 1 (2.19/0.03)
pericardial-effusion = 1: 1 (14.0)
\end{lstlisting}
Here, the existence of a pericardial effusion (fluid around the heart) is assumed to be a healthy characteristic, but it is not. Accordingly, the precision of the J48 decreases from 95\% to 88\%.

\subsubsection{Random Forest}
Let's see if a random forest in place of a simple decision tree can make better result out of the echocardiogram data set. With default settings, i.e. default number of features ($ log_2(7) + 1 = 2 $) and 10 random trees with unlimited tree depth, the random forest shows us a rather disappointing result:

\begin{lstlisting}
=== Confusion Matrix ===
  a  b   <-- classified as
  0  4 |  a = 0
  2 91 |  b = 1
\end{lstlisting}
And the statistical measures:
\begin{lstlisting}
=== Detailed Accuracy By Class ===
                 TP Rate  FP Rate  Precision  Recall  F-Measure    Class
                 0        0.022    0          0       0            0
                 0.978    1        0.958      0.978   0.968        1
Weighted Avg.    0.938    0.96     0.918      0.938   0.928     
\end{lstlisting}
\paragraph{Parameter: Number of trees}
\todo[inline]{proper line break here}
Reducing the number of the trees continuously decreases the quality of the model, as shown in the accuracy details for 5 random trees:
\begin{lstlisting}
=== Detailed Accuracy By Class ===
                 TP Rate  FP Rate  Precision  Recall  F-Measure    Class
                 0        0.032    0          0       0            0  
                 0.968    1        0.957      0.968   0.963        1   
Weighted Avg.    0.928    0.96     0.918      0.928   0.923    
\end{lstlisting}
.. and for a single random tree:
\begin{lstlisting}
=== Detailed Accuracy By Class ===
                 TP Rate  FP Rate  Precision  Recall  F-Measure    Class
                 0        0.054    0          0       0            0  
                 0.946    1        0.957      0.946   0.951        1   
Weighted Avg.    0.907    0.961    0.917      0.907   0.912    
\end{lstlisting}
Reducing the number of the trees increases the number of false positives but doesn't catch any true positives.
\paragraph{Limiting the tree-depth}
Limitations in the depths of the random trees has also almost no effect on the accuracy of random forests on this data set. The only change to the outcome can be observed as soon the depth of the trees is limited to 6:
\begin{lstlisting}
    === Confusion Matrix ===
    a  b   <-- classified as
    0  4 |  a = 0
    1 92 |  b = 1
\end{lstlisting}
Only one false positive has gone. Not really a remarkable change in the quality. This does not change if the tree-depth is further reduced.
\paragraph{Increasing the number of features}
Increasing the number of features chosen by each random tree seems to has also a low effect on the quality, because until a certain border of 5 chosen features the outcome is exactly the same as it was with default settings. But if we configure the random forests with 6 features we can spot our first true positive:
\begin{lstlisting}
=== Confusion Matrix ===
  a  b   <-- classified as
  1  3 |  a = 0
  2 91 |  b = 1
\end{lstlisting}
Which gives us the best achievable accuracy that we can reach with decision based algorithms:
\begin{lstlisting}
=== Detailed Accuracy By Class ===
                TP Rate  FP Rate  Precision  Recall  F-Measure   Class
                0.25     0.022    0.333      0.25    0.286       0
                0.978    0.75     0.968      0.978   0.973       1
Weighted Avg.   0.948    0.72     0.942      0.948   0.945      
\end{lstlisting}
\section{Adult data}

@@

\subsection{Decision Tree}
Using a decision tree yields a precision about 86\%. Varying the parameters for this algorithm influences the quality of the results only slightly, as the following examples illustrate.

First of all, with the default settings of the J48 implementation of weka, i.e. using pruning with a confidence factor of 0.25 together with MDLCorrection and Subtree raising and at least 2 instances per leaf of the tree, we get the following summary of the quality:

\begin{lstlisting}
=== Detailed Accuracy By Class ===
                TP Rate  FP Rate  Precision  Recall  F-Measure  Class
                0.615    0.062    0.759      0.615   0.68       >50K
                0.938    0.385    0.885      0.938   0.911      <=50K
Weighted Avg.   0.86     0.307    0.854      0.86    0.855      
\end{lstlisting}

The decision tree that was used for this run has a size of 413 and has 311 leaves.

\paragraph{Raising the confidence factor}
Subsequently increasing the confidence factor, which result in less pruning, leads to a continuous worse quality. If we set the confidence factor to 0.5 we get the following outcome:
\begin{lstlisting}
=== Detailed Accuracy By Class ===
                TP Rate  FP Rate  Precision  Recall  F-Measure  Class
                0.632    0.071    0.738      0.632   0.681      >50K
                0.929    0.368    0.888      0.929   0.908      <=50K
Weighted Avg.   0.857    0.296    0.852      0.857   0.854      
\end{lstlisting}
.. for a confidence factor of 0.75 we get the following:
\begin{lstlisting}
=== Detailed Accuracy By Class ===
                TP Rate  FP Rate  Precision  Recall  F-Measure  Class
                0.655    0.094    0.689      0.655   0.672      >50K
                0.906    0.345    0.892      0.906   0.899      <=50K
Weighted Avg.   0.846    0.285    0.843      0.846   0.844     
\end{lstlisting}

\paragraph{Other parameters}
Varying other parameters has a similar effect like the one mentioned above. Either the quality of the J48 algorithm decreases a little or it stays the same, thus never exceeding an overall quality of 86\%. To sum it up: Using an unpruned tree yields the same quality as a tree with a confidence factor of 0.75. Smoothing the counts at the leaves with laplace has no effect. Either without MDLcorrection or without Subtree raising during pruning, in both cases the precision falls from 86\% to 85.5\%. If we rise the minimum of instances per leaves, e.g. to 20 instances and 50 instances, the precision falls from 86\% to 85\% and 84\$ respectively.

>>>>

\missingfigure{Show the crappy null valued entries}
The only way to amend this is to add the prediction value by hand: At first, all patients who joined the survey too recently, have to be left out (i.e. survived less than one year but still alive). Then, the ``alive-at-one'' attribute has to be overwritten with the correct value (i.e. $1$ if the patient survived at least one year).
\todo{Compare results with/without preprocessing}

After preprocessing only 95 of 132 instances remain and it is still a set full of gaps with 45 missing values in total. After preprocessing it is also a very imbalanced data set for machine learning: Only 4 patients died within 12 months, a ratio of 1 to 24. As the documentation of the UCI repository suggests ``The most difficult part of this problem is correctly predicting that the patient will NOT survive. (Part of the difficulty seems to be the size of the data set.)''\cite{uci-repo}. Indeed this is rather difficult to predict with only four valid instances.

@@

\begin{lstlisting}
=== Confusion Matrix ===

  a  b   <-- classified as
  0  4 |  a = 0
  1 90 |  b = 1

@@

pericardial-effusion = 0
|   epss <= 13100: 1 (76.62/1.95)
|   epss > 13100
|   |   epss <= 15600: 0 (2.19/0.16)
|   |   epss > 15600: 1 (2.19/0.03)
pericardial-effusion = 1: 1 (14.0)
\end{lstlisting}
Here, the existence of a pericardial effusion (fluid around the heart) is assumed to be a healthy characteristic, but it is not. Accordingly, the precision of the J48 decreases from 95\% to 88\%.
\section{Adult data}

@@

\subsection{Decision Tree}
